---
title: "STAT 147 | HW"
author: "Victoria Nguyen"
subtitle: Week 2 HW
output:
  html_document:
    theme: default
    toc: yes
    toc_depth: 3
    number_sections: no
    toc_float:
      collapsed: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo = FALSE}

pre, code {
  white-space:pre !important;
  overflow-x:scroll !important}

.a-panel {
	background-color:	#e6fff9;
	padding-top: 10px;
	padding-bottom: 1px;
	padding-left: 10px;
	padding-right: 10px;
	margin-top: 10px;
	box-shadow: 0px 0px 0px 0px, 0px 6px 12px rgba(0, 0, 0, 0.2);
}

```


# Questions

## Part 1

### Q1

#### Part A

Is to be completed in Lab 2A. Below is a description of what is necessary from the lab.

- Load the following data sets in R: `ca_pop.csv`, `ca_edu.xlsx`, `ca_med_inc.sav`.

- Merge the data frames into one giant data frame.

- Creating an indicator variable for counties located in Southern California.

```{r, echo=TRUE}

#load in socal vector
socal <- c("San Luis Obispo County, California", "Kern County, California",
           "San Bernardino County, California", "Santa Barbara County, California",
           "Ventura County, California","Los Angeles County, California",
           "Orange County, California", "Riverside County, California",
           "San Diego County, California", "Imperial County, California")

```

```{r, class.output = "bg-success", class.source = "bg-success"}
install.packages('plyr', repos = "http://cran.us.r-project.org")
#loading data
#ca_pop_table2 <- read.csv("/Users/victoria/Downloads/ca_pop.csv")
#head(Pop_size)

ca_pop_table2 <- read.csv("/Users/victoria/Downloads/ca_pop.csv", 
                header= TRUE) #use header = true to import top line as headers
print(ca_pop_table2)

#install.packages("readxl")
install.packages('readxl', repos = "https://readxl.tidyverse.org/")
library("readxl")
ca_edu_table2 <- read_excel("/Users/victoria/Downloads/ca_edu.xlsx", sheet = 1)
ca_edu_table2

#install haven
install.packages('haven', repos = "http://cran.us.r-project.org")
#install.packages("haven")                             # Install haven package
library("haven")                                      # Load haven package

ca_med_inc_table <- read_sav("/Users/victoria/Downloads/ca_med_inc.sav")       # Reading data
ca_med_inc_table 

#merging as 1 frame
merged_frame1 <- merge(ca_edu_table2, ca_pop_table2 )
merged_frame2 <- merge(merged_frame1, ca_med_inc_table)
```


#### Part B

##### 1.

Create a new vector indicating how many individuals have received a Bachelor's Degree for each county.

```{r, class.output = "bg-success", class.source = "bg-success"}
sum_bach_county <- rowSums(merged_frame2[, c(2, 3, 4, 5, 6)])
sum_bach_county
```

##### 2.

How many individuals have a Bachelor's degree in Southern California?

```{r, class.output = "bg-success", class.source = "bg-success"}
#use socal subset
subset(merged_frame2, merged_frame2$num_1 > 0) 
 socal_set <- subset(merged_frame2, merged_frame2$NAME %in% socal) #creates subset using names from socal vector
 
 sum_bach_socal_county <- rowSums(socal_set[, c(2, 3, 4, 5, 6)])
 sum_bach_socal_county
   
socal_bach_sum <- sum(sum_bach_socal_county)
socal_bach_sum

```

:::{.a-panel}

5597031 bachelors degrees in southern california

::::

##### 3.

What is the average proportion of individuals in Southern California who have a Bachelor's Degree?

```{r, class.output = "bg-success", class.source = "bg-success"}

# find number of individuals per county in socal
#creates vector of just county pops for socal
socal_county_pop <- socal_set[, c(7)]
socal_county_pop
# find proportion of individuals with bach in socal 
#creates vector of the proportions
socal_county_bach_prop <-    (sum_bach_socal_county / socal_county_pop)
socal_county_bach_prop

mean(socal_county_bach_prop)

#the average socal county has a proportion of 0.2126916 bachelors degree owners
```

:::{.a-panel}
the average socal county has a proportion of 0.2126916 bachelors degree owners
::::

##### 4.

The correlation measures the association between numeric variables. It can have a number between -1 and 1. The closer the value is to 0, the stronger evidence there is no association. The closer the value is to 1/-1, the stronger evidence of a positive/negative association. 

You can find the correlation between two vectors using the `cor` function. You will only need to specify the two variables separated by a comma in the function.

Find the correlation between median income and Proportion of Bachelor's Degree in each county.

```{r, class.output = "bg-success", class.source = "bg-success"}
#vector of median income
cal_median_inc <- merged_frame2[, c(8)]
cal_median_inc

#vector of bach degrees
sum_bach_cal_county <- rowSums(merged_frame2[, c(2, 3, 4, 5, 6)])
sum_bach_cal_county

#vector of pops
cal_county_pop <- merged_frame2[, c(7)]
cal_county_pop

#proportions of degrees per county
cal_bach_prop <- (sum_bach_cal_county / cal_county_pop)

cor(cal_bach_prop, cal_median_inc)

```

:::{.a-panel}

correlation between median income and Proportion of Bachelor's Degree in each county: 0.141871

::::


##### 5.

What is the correlation between median income and proportion of Bachelor's degree in Southern California.

```{r, class.output = "bg-success", class.source = "bg-success"}
#vector of median income in socal
socal_county_median_inc <- socal_set[, c(8)]
socal_county_median_inc

cor(socal_county_bach_prop, socal_county_median_inc)

```

:::{.a-panel}
correlation between median income and Proportion of Bachelor's Degree in each county: 0.8832506

::::

## Part 2

### Q2

Recreate the matrices below.  Please use the same names as those shown.

```{r, class.output = "bg-success", class.source = "bg-success"}
 #sqr_even_mat
sqr_even_mat.data <- c(2,4,6,8,10,12,14,16,18,20,22,24,26,28,30,32) #inputting the values
sqr_even_mat <- matrix(sqr_even_mat.data,nrow=4,ncol=4,byrow=TRUE)
sqr_even_mat

#two_col_mat
two_col_mat.data <- c(8,6,16,14,24,22,32,30) #inputting the values
two_col_mat <- matrix(two_col_mat.data,nrow=4,ncol=2,byrow=TRUE)
two_col_mat

```

### Q3

Multiply the two matrices `two_col_mat` and `sqr_even_mat`.  Store the result in a variable named `prod_mat`.  Print the matrix to check.

```{r, class.output = "bg-success", class.source = "bg-success"}
prod_mat <- sqr_even_mat %*% two_col_mat
prod_mat
```

### Q4

Let's return to the two original matrices, `sqr_even_mat` and `two_col_mat`. Bind the columns together, resulting in a matrix with 6 columns.  Place `sqr_even_mat` on the left and `two_col_mat` on the right. Store the result as `six_col_mat`

```{r, class.output = "bg-success", class.source = "bg-success"}
six_col_mat <- cbind(sqr_even_mat,two_col_mat)
six_col_mat

```

### Q5

Next we want to bind the rows.  What is wrong with the following code?

```{r, class.output = "bg-success", class.source = "bg-success"}
#rbind(sqr_even_mat, two_col_mat)
#we cannot combine the two as there needs to be an equal number of columns to combine two matrices by row
#we need to use the plyr package in order to combine two matrices with different columns
install.packages('plyr', repos = "http://cran.us.r-project.org")
plyr::rbind.fill.matrix(sqr_even_mat, two_col_mat)

```

:::{.a-panel}
a package is needed to fill the "empty columns" with NA if we want to bind matrices by row with uneven column numbers. 
::::

## Part 3

The next questions will involve the `penguins` data set from the `palmerpenguins` package.

```{r, class.output = "bg-success", class.source = "bg-success"}
#install.packages("palmerpenguins")
install.packages('palmerpenguins', repos = "https://cran.r-project.o")
library(palmerpenguins)
#eliminate observations with missing values
new_penguin <- na.omit(penguins)
new_penguin

```

### Q6

For each penguin species, find the mean `bill_length_mm`. Remember to eliminate the missing observations.

```{r, class.output = "bg-success", class.source = "bg-success"}

#Adelie
Adelie_set <- subset(new_penguin, new_penguin$species %in% "Adelie") #create subset
Adelie_bill_length_set <- subset(Adelie_set, Adelie_set$bill_length_mm >0) #create subset of just bill lengths, removes empty values
Adelie_BL_mean <- mean(Adelie_bill_length_set$bill_length_mm)
Adelie_BL_mean

#Gentoo
Gentoo_set <- subset(new_penguin, new_penguin$species %in% "Gentoo") #create subset
Gentoo_bill_length_set <- subset(Gentoo_set, Gentoo_set$bill_length_mm >0) #create subset of just bill lengths, removes empty values
Gentoo_BL_mean <- mean(Gentoo_bill_length_set$bill_length_mm)
Gentoo_BL_mean

#Chinstrap
Chinstrap_set <- subset(new_penguin, new_penguin$species %in% "Chinstrap") #create subset
Chinstrap_bill_length_set <- subset(Chinstrap_set, Chinstrap_set$bill_length_mm >0) #create subset of just bill lengths, removes empty values
Chinstrap_BL_mean <- mean(Chinstrap_bill_length_set$bill_length_mm)
Chinstrap_BL_mean
```

:::{.a-panel}

Adelie: 38.82397
Gentoo: 47.56807
Chinstrap: 48.83382

::::

### Q7

For each species and sex, find the mean `bill_length_mm`.

```{r, class.output = "bg-success", class.source = "bg-success"}
#Adelie
Adelie_F_bill_length_set <- subset(Adelie_bill_length_set, Adelie_bill_length_set$sex %in% 
                                     "female") 
Adelie_M_bill_length_set <- subset(Adelie_bill_length_set, Adelie_bill_length_set$sex %in% 
                                     "male") 
Adelie_F_BL_mean <- mean(Adelie_F_bill_length_set$bill_length_mm)
Adelie_M_BL_mean <- mean(Adelie_M_bill_length_set$bill_length_mm)
Adelie_F_BL_mean 
Adelie_M_BL_mean 

#Gentoo
Gentoo_F_bill_length_set <- subset(Gentoo_bill_length_set, Gentoo_bill_length_set$sex %in% 
                                     "female") 
Gentoo_M_bill_length_set <- subset(Gentoo_bill_length_set, Gentoo_bill_length_set$sex %in% 
                                     "male") 
Gentoo_F_BL_mean <- mean(Gentoo_F_bill_length_set$bill_length_mm)
Gentoo_M_BL_mean <- mean(Gentoo_M_bill_length_set$bill_length_mm)
Gentoo_F_BL_mean 
Gentoo_M_BL_mean 

#Chinstrap
Chinstrap_F_bill_length_set <- subset(Chinstrap_bill_length_set, Chinstrap_bill_length_set$sex %in% 
                                     "female") 
Chinstrap_M_bill_length_set <- subset(Chinstrap_bill_length_set, Chinstrap_bill_length_set$sex %in% 
                                     "male") 
Chinstrap_F_BL_mean <- mean(Chinstrap_F_bill_length_set$bill_length_mm)
Chinstrap_M_BL_mean <- mean(Chinstrap_M_bill_length_set$bill_length_mm)
Chinstrap_F_BL_mean 
Chinstrap_M_BL_mean 

```

:::{.a-panel}
Adelie female: 37.25753
Adelie male: 40.39041
Gentoo female: 45.56379
Gentoo male: 49.47377
Chinstrap female: 46.57353
Chinstrap male: 51.09412

::::

### Q8

Create a two new variables indicating if the penguin is in the top $50^{th}$ percentile (including $50^{th}$) for the variables `bill_length_mm` and `bill_depth_mm` . Generate a $2\times 2$ contingency table.

```{r, class.output = "bg-success", class.source = "bg-success"}

# 50th percentile for bill length
bill_length_50 <- median(new_penguin$bill_length_mm)

# 50th percentile for bill depth
bill_depth_50 <- median(new_penguin$bill_depth_mm)

#there is a better way to do this but i have a smooth brain
#making a new vector so i don't accidentally ruin new_penguin
new_penguin2 <-data.frame(new_penguin)

#creates a column giving TRUE if the penguin's bill length is greater than or equal to the median (meaning it is in the top 50th percentile)
new_penguin2$BL_in_50_percentile <- (new_penguin2$bill_length_mm >= bill_length_50)

#creates a column giving TRUE if the penguin's bill depth is in the top 50th percentile
new_penguin2$BD_in_50_percentile <- (new_penguin2$bill_depth_mm >= bill_depth_50)

# 2x2 contingency table ... length & depth vs in 50th percentile vs not


#convert results to  data frame
BL_BD_50_table <- table(new_penguin2$BD_in_50_percentile, new_penguin2$BL_in_50_percentile)
  print.table(BL_BD_50_table) # doesn't have good labels, adding them so table is better for viewing
  
dimnames(BL_BD_50_table) <- list("Bill Depth" = c("<50", "50≤"),
                                 "Bill Length" = c("<50", "50≤"))

BL_BD_50_table

# checking using filters.. 
# 52 = both below 
# 109 = BL in top 50, BD below
# 114 = BL below, BD above
# 58 = both in top 50
```


### Q9

Using the new variables, find the mean `body_mass_g` for each $50^{th}$ percentile indicator variable `bill_length_mm` and `bill_depth_mm`.

```{r, class.output = "bg-success", class.source = "bg-success"}
# does this mean find the mean body mass above and below for each?
# using this as a guide https://stackoverflow.com/questions/29724286/calculate-a-mean-by-criteria-in-r


#install.packages("dplyr")
install.packages('dplyr', repos = "http://cran.us.r-project.org")
library(dplyr)
new_penguin2  %>% group_by(BL_in_50_percentile) %>%
  summarise(mean=mean(body_mass_g))

new_penguin2  %>% group_by(BD_in_50_percentile) %>%
  summarise(mean=mean(body_mass_g))

new_penguin2  %>% group_by(BL_in_50_percentile, BD_in_50_percentile) %>% 
  summarise(mean=mean(body_mass_g))


```

:::{.a-panel}
Mean Body Mass Average for:

Beak Length in top 50th Percentile: 4639.072
Beak Length Below 50th Percentile: 3772.440	
Beak Depth in top 50th Percentile: 3835.756
Beak Depth Below 50th Percentile: 4603.727		

BL Below BD Below:	3683.173		
BL Below BD Top 50:	3813.158		
BL Top 50	BD Below:	5042.890		
BL Top 50	BD Top 50:	3880.172	

(at this point i'm pretty sure i overthought the question)
::::

### Q10

Find the ratio of the standardized `bill_length_mm` and standardized `bill_depth_mm` for each island: $ std(bill_length_mm) / std(bill_depth_mm)$. To standardize a variable, use the z-score formula: 
$$
z = \frac{x-mean(x)}{sd(x)}.
$$

```{r, class.output = "bg-success", class.source = "bg-success"}
#Biscoe
Biscoe_set <- subset(new_penguin2, new_penguin2$island %in% "Biscoe") #create subset
##BL
Biscoe_BL_mean <- mean(Biscoe_set$bill_length_mm)
Biscoe_BL_sd <- sd(Biscoe_set$bill_length_mm)
##BD
Biscoe_BD_mean <- mean(Biscoe_set$bill_depth_mm)
Biscoe_BD_sd <- sd(Biscoe_set$bill_depth_mm)
##Standardizing values
Biscoe_std_BL <- (( Biscoe_set$bill_length_mm  - Biscoe_BL_mean ) / Biscoe_BL_sd)
Biscoe_std_BD <- (( Biscoe_set$bill_depth_mm  - Biscoe_BD_mean ) / Biscoe_BD_sd)
#used these to check, mean should be 0 or very close. sd should be 1. 
mean(Biscoe_std_BL) 
sd(Biscoe_std_BL)
##Ratio
Biscoe_std_BLBD_ratio <- ( Biscoe_std_BL / Biscoe_std_BD )
##Mean Ratio
Biscoe_mean_ratio <- mean(Biscoe_std_BLBD_ratio)
Biscoe_mean_ratio


#Dream
Dream_set <- subset(new_penguin2, new_penguin2$island %in% "Dream") #create subset
##BL
Dream_BL_mean <- mean(Dream_set$bill_length_mm)
Dream_BL_sd <- sd(Dream_set$bill_length_mm)
##BD
Dream_BD_mean <- mean(Dream_set$bill_depth_mm)
Dream_BD_sd <- sd(Dream_set$bill_depth_mm)
##Standardizing values
Dream_std_BL <- (( Dream_set$bill_length_mm  - Dream_BL_mean ) / Dream_BL_sd)
Dream_std_BD <- (( Dream_set$bill_depth_mm  - Dream_BD_mean ) / Dream_BD_sd)
##Ratio
Dream_std_BLBD_ratio <- ( Dream_std_BL / Dream_std_BD )
##Mean Ratio
Dream_mean_ratio <- mean(Dream_std_BLBD_ratio)
Dream_mean_ratio

#Torgersen
Torgersen_set <- subset(new_penguin2, new_penguin2$island %in% "Torgersen") #create subset
##BL
Torgersen_BL_mean <- mean(Biscoe_set$bill_length_mm)
Torgersen_BL_sd <- sd(Biscoe_set$bill_length_mm)
##BD
Torgersen_BD_mean <- mean(Torgersen_set$bill_depth_mm)
Torgersen_BD_sd <- sd(Torgersen_set$bill_depth_mm)
##Standardizing values
Torgersen_std_BL <- (( Torgersen_set$bill_length_mm  - Torgersen_BL_mean ) / Torgersen_BL_sd)
Torgersen_std_BD <- (( Torgersen_set$bill_depth_mm  - Torgersen_BD_mean ) / Torgersen_BD_sd)
##Ratio
Torgersen_std_BLBD_ratio <- ( Torgersen_std_BL / Torgersen_std_BD )
##Mean Ratio
Torgersen_mean_ratio <- mean(Torgersen_std_BLBD_ratio)
Torgersen_mean_ratio
```

:::{.a-panel}
Mean Ratios: 
Biscoe: -6.860314
Dream: -0.3108084
Torgersen: 1.001905
::::

## Part 4

### Q11

The `readr` package has two separate functions to read data that are separated by white space (space, tab, etc...): `read_table` and `read_table2`. Read the help documentation for the functions and explain why you may want to use one function over the other function.

```{r, class.output = "bg-success", class.source = "bg-success"}
#install.packages("readr")
install.packages('readr', repos = "http://cran.us.r-project.org")
library("readr")
help("readr")
```

:::{.a-panel}

One would want to use read_table2() when they would like to customize the length of lines and 
the width of the white space between columns. 

read_table() is better when the user would like the white space and line length to be chosen for them automatically. this may be useful when there are multiple tables being read and the user wants the presentation to be as consistent as possible. 

::::

### Q12

Both `lapply` and `sapply` functions belong to the same family of `*apply` functions where a user specified function is applied to an R object. What is the difference between the `sapply` and `lapply` function.

:::{.a-panel}

lapply returns a LIST with the same/similar length as the original list object (list, vector, or data frame) that has been inputted.

sapply is similar to lapply but returns a VECTOR or MATRIX. 

::::

### Q13

In linear algebra, the Kronecker product creates block matrices from two matrices. The Kronecker product multiplies each element of the first matrix the entire second matrix. What is the function in R to find the Kronecker product?

:::{.a-panel}
kronecker(X, Y, FUN = "*", make.dimnames = FALSE)
X %x% Y

::::

### Q14

What function do you use to view what files are in your working directory?

:::{.a-panel}
list.files()
::::

### Q15

Fix the code.

```{r, class.output = "bg-success", class.source = "bg-success"}
#install package
install.packages('palmerpenguins', repos = "https://cran.r-project.o")

#load package
library(palmerpenguins)

tapply(new_penguin$bill_length_mm, list(new_penguin$island), mean)
#needed to use the new_penguins data frame, which removed rows with empty cells. 
```
