---
title: 'STAT 147 | HW'
subtitle: 'Week 3 HW'
author: "Victoria Nguyen"
output:
  html_document:
    theme: default
    toc: true
    toc_depth: 3
    number_sections: false
    toc_float:
      collapsed: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

```{css, echo = FALSE}
pre, code {
  white-space:pre !important;
  overflow-x:scroll !important}

.a-panel {
	background-color: rgba(49, 126, 172, 0.1);
	padding-top: 10px;
	padding-bottom: 1px;
	padding-left: 10px;
	padding-right: 10px;
	margin-top: 10px;
	box-shadow: 0px 0px 0px 0px, 0px 6px 12px rgba(0, 0, 0, 0.2);
	border-bottom: 5px solid #317EAC;
}

```

------------------------------------------------------------------------

# **Instructions**

The goal of this assignment is to continue our exploration of loops, while loops, and if and else statements.  Please use the RMD file `FIRST_LAST_HW_03.Rmd` to answer your questions. Delete code or answer chunks as necessary. You will submit both the RMD file and the corresponding HTML file from the knitted document.

Note: you may need to search the internet to find answers for some of the questions.

You will be graded as follows:

-   Does your R chunks run (some errors are acceptable in this assignment)?
-   Have you completed the assignment in its entirety?
-   Have you followed the instructions carefully?
-   Have you responded to the questions correctly?

------------------------------------------------------------------------

#### Grading

-   **(20 pts)** Have you completed the assignment in its entirety?

-   **(20 pts)** Are your responses correct for the subset of randomly graded questions?

-   **(5 pts)** CODE DOCUMENTATION. Add comments to nearly every line explaining what the line of code is doing.

-   **(5 pts)** Submitted Knitted Document (HTML File)

------------------------------------------------------------------------

# Questions

## Part 1

### Q1

#### Part A

Is to be completed in Lab 3B. Below is a description of what is necessary from the lab.

- Generating a vector filled with Fibonacci Numbers.

- Generating a vector filled with Triangular Numbers.

- Generating a vector filled with Tetrahedral Numbers.


```{r, class.output = "bg-success", class.source = "bg-success"}

#Include Code HERE from Lab 3B

```



#### Part B

##### 1.

Create a vector containing only the even numbers from the Fibonacci sequence. Hint: Use loops, if else statements and divide each number by 2 using `%%`.
Find the sum of all the even Fibonacci numbers?

```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


##### 2.

Create a vector containing only the even numbers from the Triangular sequence. Hint: Use loops, if else statements and divide each number by 2 using `%%`.
Find the sum of all the even Triangular numbers?

```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


##### 3.

Create a vector containing only the even numbers from the Tetrahedral sequence. Hint: Use loops, if else statements and divide each number by 2 using `%%`.
Find the sum of all the even Tetrahedral numbers?

```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


## Part 2

Use the following matrix for Part 2 questions:

```{r, echo=TRUE}
x <- matrix(1:100, nrow = 10)
x
```

### Q2

The `colSums` function will add all the values in each column of a matrix. Instead of using the `colSums` function, write a loop that will produce the same results.


```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


### Q3

The `rowMeans` function will compute the mean for each row in a matrix. Instead of using the `rowMeans` function, write a loop that will produce the same results.


```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


### Q4

Create a logical $10 \times 10$ matrix indicating whether each element in matrix `x` is either even  (TRUE) or odd (FALSE).

```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


## Part 3

### Q5

Write a for loop that iterates over the column names of the `mtcars` data set and prints the name with the number of characters in the column name in parentheses.

Example output: `"mpg (3)"`.

Use the following functions, and any others you wish: `paste` and `nchar` functions.

Note: `nchar` function counts how many characters are in a string.

```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


### Q6

Use a while loop to investigate the number of terms required until the product of

1⋅2⋅3⋅4⋅...

reaches above 1 million. How many elements is needed to reach 1 million.

```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


### Q7

Write a while loop that prints out standard random normal numbers (use `rnorm`) but stops (breaks) if you get a number bigger than 1.

```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


### Q8

Using a while loop, generate a 100 positive values from a standard normal distribution. Your vector should look similar to the vector below.

```{r , echo = FALSE}
abs(rnorm(100))
```

```{r, class.output = "bg-success", class.source = "bg-success"}
## Type Code Here
## Delete chunk if necessary
```

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


## Part 4

### Q9

Watch the following video [https://www.youtube.com/watch?v=Mv9NEXX1VHc](https://www.youtube.com/watch?v=Mv9NEXX1VHc):


What is recursion?

::: {.a-panel}
Type Response Here.

Delete if necessary.
:::


### Q10

Watch the following video [https://www.youtube.com/watch?v=HXNhEYqFo0o](https://www.youtube.com/watch?v=HXNhEYqFo0o):


What is the difference between loops and recursion?


::: {.a-panel}
Type Response Here.

Delete if necessary.
:::

