---
title: 'STAT 147 | HW'
subtitle: 'Week 3 HW'
author: "Victoria Nguyen"
output:
  html_document:
    theme: default
    toc: true
    toc_depth: 3
    number_sections: false
    toc_float:
      collapsed: true
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{css, echo = FALSE}
pre, code {
  white-space:pre !important;
  overflow-x:scroll !important}

.a-panel {
	background-color: rgba(49, 126, 172, 0.1);
	padding-top: 10px;
	padding-bottom: 1px;
	padding-left: 10px;
	padding-right: 10px;
	margin-top: 10px;
	box-shadow: 0px 0px 0px 0px, 0px 6px 12px rgba(0, 0, 0, 0.2);
	border-bottom: 5px solid #317EAC;
}

```

------------------------------------------------------------------------

# **Instructions**

The goal of this assignment is to continue our exploration of loops, while loops, and if and else statements.  Please use the RMD file `FIRST_LAST_HW_03.Rmd` to answer your questions. Delete code or answer chunks as necessary. You will submit both the RMD file and the corresponding HTML file from the knitted document.

Note: you may need to search the internet to find answers for some of the questions.

You will be graded as follows:

-   Does your R chunks run (some errors are acceptable in this assignment)?
-   Have you completed the assignment in its entirety?
-   Have you followed the instructions carefully?
-   Have you responded to the questions correctly?

------------------------------------------------------------------------

#### Grading

-   **(20 pts)** Have you completed the assignment in its entirety?

-   **(20 pts)** Are your responses correct for the subset of randomly graded questions?

-   **(5 pts)** CODE DOCUMENTATION. Add comments to nearly every line explaining what the line of code is doing.

-   **(5 pts)** Submitted Knitted Document (HTML File)

------------------------------------------------------------------------

# Questions

## Part 1

### Q1

#### Part A

Is to be completed in Lab 3B. Below is a description of what is necessary from the lab.

- Generating a vector filled with Fibonacci Numbers.

- Generating a vector filled with Triangular Numbers.

- Generating a vector filled with Tetrahedral Numbers.


```{r, class.output = "bg-success", class.source = "bg-success"}

#Fibonacci
fibonacci <- c(0,1)               #initialize the first two numbers: 0 & 1 in a 
for(i in (3:50)){                 #for loop for the 3rd to 50th numbers in the sequence
  fibonacci[i] <- (fibonacci[i-1]+fibonacci[i-2]) #each number is found using the two before it
}
fibonacci 

#Triangular

Triangular <- c(1) 
for (i in (2:40))
  Triangular[i] <- (i * (i + 1 )) / 2
  
#Returns as a vector
  Triangular

#Tetrahedral

x <- 1
z <- c(1:39)
tetrahedral <- c()
for (i in seq_along(z)){
  tetrahedral[i] <- (x * (x + 1) * (x + 2)) /6
  x <- x + 1
}
tetrahedral
```



#### Part B

##### 1.

Create a vector containing only the even numbers from the Fibonacci sequence. Hint: Use loops, if else statements and divide each number by 2 using `%%`.
Find the sum of all the even Fibonacci numbers?

```{r, class.output = "bg-success", class.source = "bg-success"}

fibonacci_even <- c() #initialize empty vector
n <- 1;
for (i in (1:50) ) {
    x <- 0
    #even
    if (fibonacci[n]%%2==0)
    {
        x<-fibonacci[n]
    }
    #odd
    if (x > 0)
        fibonacci_even<-c(fibonacci_even, x);
        n <- n + 1;
        
}
fibonacci_even

sum(fibonacci_even)
```

::: {.a-panel}
Sum of the even fibonacci numbers from the first 50: 6293134512
:::


##### 2.

Create a vector containing only the even numbers from the Triangular sequence. Hint: Use loops, if else statements and divide each number by 2 using `%%`.
Find the sum of all the even Triangular numbers?

```{r, class.output = "bg-success", class.source = "bg-success"}


Tri_even <- c()
n <- 1;
for (i in (1:40) ) {
    x <- 0
    #even
    if (Triangular[n]%%2==0)
    {
        x<-Triangular[n]
    }
    if (x != 0)
        Tri_even<-c(Tri_even, x);
        n <- n + 1;
        
}
Tri_even
sum(Tri_even)
```

::: {.a-panel}
Sum of the even triangular numbers: 6160
:::


##### 3.

Create a vector containing only the even numbers from the Tetrahedral sequence. Hint: Use loops, if else statements and divide each number by 2 using `%%`.
Find the sum of all the even Tetrahedral numbers?

```{r, class.output = "bg-success", class.source = "bg-success"}

tetrahedral_even <- c()
n <- 1;
for (i in (1:39) ) {
    x <- 0
    #even
    if (tetrahedral[n]%%2==0)
    {
        x<-tetrahedral[n]
    }
    if (x != 0)
        tetrahedral_even<-c(tetrahedral_even, x);
        n <- n + 1;
        
}
tetrahedral_even
sum(tetrahedral_even)
```

::: {.a-panel}
sum of even tetrahedral numbers = 85430
:::


## Part 2

Use the following matrix for Part 2 questions:

```{r, echo=TRUE}
x <- matrix(1:100, nrow = 10)
x
```

### Q2

The `colSums` function will add all the values in each column of a matrix. Instead of using the `colSums` function, write a loop that will produce the same results.


```{r, class.output = "bg-success", class.source = "bg-success"}

#initialize 10 x 10 matrix
x <- matrix(1:100, nrow = 10)

#loop for sums across 10 columns
colSums_loop <- vector(mode = "numeric")
for(i in 1:10) {
  colSums_loop[i] <- sum(x[1:10, i])
}
colSums_loop

```


### Q3

The `rowMeans` function will compute the mean for each row in a matrix. Instead of using the `rowMeans` function, write a loop that will produce the same results.


```{r, class.output = "bg-success", class.source = "bg-success"}
#get mean across 10 rows, uses same x as previous question
rowMeans_loop <- vector(mode = "numeric")
for(i in 1:10) {
  rowMeans_loop[i] <- mean(x[i, 1:10])
}
rowMeans_loop
```



### Q4

Create a logical $10 \times 10$ matrix indicating whether each element in matrix `x` is either even  (TRUE) or odd (FALSE).

```{r, class.output = "bg-success", class.source = "bg-success"}
x <- matrix(1:100, nrow = 10)
x

#empty 10 x 10 matrix
x_logical <- matrix(nrow = 10, ncol = 10 ) 
n <- 1
for (i in seq_along(x) ) {
    num <- x[i]
    #even
    if (num%%2==0)
    {
        x_logical[n] = "even"
        n=n+1
    }
    #odd
    else
    {
      x_logical[n] = "odd"
      n=n+1
    }
        
}

x_logical

```


## Part 3

### Q5

Write a for loop that iterates over the column names of the `mtcars` data set and prints the name with the number of characters in the column name in parentheses.

Example output: `"mpg (3)"`.

Use the following functions, and any others you wish: `paste` and `nchar` functions.

Note: `nchar` function counts how many characters are in a string.

```{r, class.output = "bg-success", class.source = "bg-success"}
data(mtcars)

#initialize the two variables
col_names <- colnames(mtcars) # listn of all col names
names_count <- nchar(col_names) #counts chats in col name

#goes down col_names, pastes name then character count
for (i in seq_along(col_names)){
    print(paste(col_names[i], " (",(names_count[i]), ")",sep = ""))

}

```

### Q6

Use a while loop to investigate the number of terms required until the product of

1⋅2⋅3⋅4⋅...

reaches above 1 million. How many elements is needed to reach 1 million.

```{r, class.output = "bg-success", class.source = "bg-success"}
#initialize beginning values, gets factorial of 1
product <- 0
a <- 1
b <- 1
for (i in b)
while (product <= 1000000) #keeps going until it hits the first product above 1,000,000
  {
      product[i] <- factorial(a+b-1)/factorial(a-1)
      
      #adds element if product hasn't reached 1,000,000
      if (product <= 1000000) 
      b <- b + 1
  }
product
b
```

::: {.a-panel}
10 elements are needed to reach the first value greater than or equal to 1 million. The first product after breaking 1,000,000 is 3,628,800
:::


### Q7

Write a while loop that prints out standard random normal numbers (use `rnorm`) but stops (breaks) if you get a number bigger than 1.

```{r, class.output = "bg-success", class.source = "bg-success"}

set.seed(245424)

r <- 0
while(r<=1) #stops at first value above 1
{
  r <- rnorm (n = 1)
  print(r) 
}
```


### Q8

Using a while loop, generate a 100 positive values from a standard normal distribution. Your vector should look similar to the vector below.

```{r , echo = FALSE}

abs(rnorm(100))

```

```{r, class.output = "bg-success", class.source = "bg-success"}

pos_100 <- c() #initialize empty vector
i <- 1
while (i <= 100) #limit to 100 
{
  p = rnorm(1,0,1)
  if (p>0) #only keep positive values
  {
    pos_100[i] <- p
    i = i + 1
  }
}
pos_100
```



## Part 4

### Q9

Watch the following video [https://www.youtube.com/watch?v=Mv9NEXX1VHc](https://www.youtube.com/watch?v=Mv9NEXX1VHc):


What is recursion?

::: {.a-panel}
Recursion is when data can be passed through a function and passed through itself until conditions for termination (base condition) are met, in which all data up until putting it out will be returned. 
:::


### Q10

Watch the following video [https://www.youtube.com/watch?v=HXNhEYqFo0o](https://www.youtube.com/watch?v=HXNhEYqFo0o):


What is the difference between loops and recursion?


::: {.a-panel}
A loop will continue with the action until it reaches the condition for "ending the list", meanwhile recursion will return data to the function if the base condition to terminate isn't met. 
:::

