---
title: "Inference with Categorical Data"
author: "Victoria Nguyen"
date: "7/18/2021"
output: 
  html_document:
    toc: true
    toc_float:
      toc_collapsed: true
    toc_depth: 3
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = FALSE)
```

```{css, echo = FALSE}
.o-panel {
	background-color: #E8F8F5 ;
	padding-top: 10px;
	padding-bottom: 1px;
	padding-left: 10px;
	padding-right: 10px;
	margin-top: 10px;
}

.c-panel {
	background-color: #EBF5FB;
	padding-top: 10px;
	padding-bottom: 1px;
	padding-left: 10px;
	padding-right: 10px;
	margin-top: 10px;
}


```

## Introduction

This tutorial focuses on inferences with categorical data in R. 

### Contents
1.  What is Inference?

2.  One-Sample Proportion Test

3.  Two-Sample Proportion Test

## What is Inference?

**Statistical inference** allows us to make deduce information about a sample or population's properties. It's most commonly done by calculating estimates or by testing hypotheses. 


## One-Sample Proportion Test

Given a sample of data from a population, we can use R to compare the porportions of successes and failures to the true proportion, `p`, of the population using `prop.test(x,n,p)`, where `n` is the sample size, `x` is the number of successes, and `p` is the true proportion.

The output will be a p-value. If it's less than 0.05, it's unlikely that the sample proportion is the same as `p`. A p-value above 0.05 is insufficient to drawing such conclusions. 


### Example

Suppose my friend assumed the Bucks would win 4 out of 7 games needed to win the finals. We know they played 72 games and won 46 of them. In this scenario, all 7 games are played. $\frac{4}{7}$ translates to roughly $.57$.

#### Testing the sample proportion 

```{r Inf_1, echo = TRUE}
#here, n=72, x=46, and .57 is the proportion we're comparing to. We use "greater" here as we are looking for 4 or more games out of the 7. 

prop.test(46, 72, 0.57, alternative = "greater")

```
Notice that the above result includes the 95% confidence interval. It's set to .95 by default as that is the most commonly observed, however we can change it like this:

```{r Inf_2, echo = TRUE}
prop.test(46, 72, 0.57, conf.level = .99, alternative = "greater")
```
## Two-Sample Proportion Test

Suppose that, instead of comparing a sample proportion to the population, we wanted to compare two samples for equal proportions. 


### Example

Suppose basketball teams stayed the same between this season and last, however the Minnesota Timberwolves changed coaches. For the 2020-2021 season (Coach B), they went 23-49, and 19-45 for the 2019-2020 season (Coach A). 

Compare the two seasons to see if the change to Coach B from Coach A made a significant difference in team performance. 

#### Comparing the Sample Proportions

First, let's create the arrays we need for wins and games.

```{r Inf_3, echo = TRUE}
wins <- c(19, 23)
games <- c(19+45, 23+49)
```

Now, let's perform a `prop.test()`:

```{r Inf_4, echo = TRUE}
prop.test(wins, games)
```
::: {.o-panel}
The high p-value of 0.9216 shows that we cannot reject the null hypothesis: that Coach B's performance caused a significant increase in the Minnesota Timberwolves' success. 
:::

## Resources

Here are other resources for Categorical Inference in R:

1.  [Quick-R](https://rc2e.com/generalstatistics#recipe-id134)
2.  [DWR44 (YouTube): Two Porportion Z-Test in R](https://www.youtube.com/watch?v=xWwnVYO4Aus)
3.  [DWR44 (YouTube): 1-Proportion Z-Test in R](https://www.youtube.com/watch?v=7vmdHIG9NR8)
4.  [Bookdown: Chi-square Test in R](https://bookdown.org/ndphillips/YaRrr/chi-square-chsq-test.html)

